---
title: "The Preston Curve Revisited"
author: "Joachim Gassen"
institute: "TRR 266 Accounting for Transparency"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  beamer_presentation: 
    latex_engine: xelatex # pdflatex creates rastered fonts
    slide_level: 3

        
header-includes:
- \usepackage{booktabs} 
- \usepackage{threeparttable}
- \usepackage{graphicx}
---


```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(
  cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE
)
library(tidyverse)
library(kableExtra)
library(GGally)
library(corrplot)
load("../output/results.rda")
load("../data/generated/sample.rda")
```


# Motivation

### The Preston Curve

\begin{center}
\includegraphics[width=0.8\textwidth]{../media/preston_1975_p235.png} \\
Preston (1975, p. 235)
\end{center}


### Data

```{r tab_descriptives_sample, results="asis"}
desc_rnames <- paste0("\\textit{", smp_def$label[7:9], "}")
t <- tab_desc_stat
rownames(t$df) <- desc_rnames
names(t$df)[c(5,7)] <- c("25 \\%", "75 \\%")
kable(
  t$df, digits = c(0, 3, 3, 3, 3, 3, 3, 3), format = "latex",
  caption = "\\label{tab:descriptives}Descriptive Statistics",
  format.args = list(decimal.mark = ".", big.mark = ",", 
                     scientific=FALSE),
  booktabs = TRUE, escape = FALSE, linesep = ""
) -> kab_latex

lat_tab <- unlist(strsplit(kab_latex, "\n"))
lat_tab[6] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
lat_tab[8] <- "\\hline\\\\[-1.8ex]"
lat_tab[length(lat_tab) - 2] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"

latex_tab <- c(
  "\\begin{table}[!htbp]",  
  "\\resizebox{\\textwidth}{!}{",
  "\\begin{threeparttable}",
  lat_tab[5:(length(lat_tab) - 2)],
  "\\end{tabular}",
  "\\begin{tablenotes}[flushleft]",
  "\\setlength{\\labelsep}{0pt}",
  sprintf(
    "\\item Note: The data is obtained from the World Bank. The sample covers %d countries and the period %d to %d. \\textit{National Income per captia} values are in constant 2010 thousand U.S. dollars.", length(unique(smp$country)),
    min(as.numeric(as.character(smp$year))), 
    max(as.numeric(as.character(smp$year)))
  ),
 "\\end{tablenotes}",
  "\\end{threeparttable}",
  "}",
  "\\end{table}"
)
cat(paste(latex_tab, collapse = "\n"))  
```

### Regional Descriptives

```{r reg_desc, echo=F, results="asis"}
kbl(regional_descripives, booktabs = T) %>%
  kable_styling(latex_options = c("scale_down", "hold_position"))
```

### Life Expectancy by National Income

```{r Scatter}
fig_scatter
```

### No presentation without correlation table...

``` {r CorrTable, results="asis"}
tab <- tab_corr
lat_tab <- unlist(strsplit(tab$kable_ret, "\n"))
lat_tab[7:9] <- str_replace(
  lat_tab[7:9], fixed(smp_def$var_name[8:10]), smp_def$label[8:10]
)
lat_tab[4] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
lat_tab[6] <- "\\hline\\\\[-1.8ex]"
lat_tab[10] <- "\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]"
latex_tab <- c(
  "\\begin{center}",
  "\\begin{tabular}{lccc}",
  lat_tab[4:10],
  "\\end{tabular}",
  "\\end{center}"
)
cat(paste(latex_tab, collapse = "\n"))
```

### Pearson Correlations by Region in a Table...

```{r reg_cor, echo=F}
kbl(regional_correlations, booktabs = T) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  row_spec(7, bold = T, color = "red")
```

### ... and displayed in a Graph

```{r Corr, echo = FALSE, fig.align="center", fig.cap="\\label{fig:reg_cor_graph}Pearson Correlations Regional Differences", fig.width=8, fig.height=8}
subset <- smp %>% select(ln_gdp_capita, life_expectancy, unemployment, region)
# ggpairs(subset, columns = 1:3 ,ggplot2::aes(colour = region), upper = list(continuous = wrap("cor", size = 3)))

colnames(subset) <- c("ln(National Income \n per Capita)", "Life Expectancy \n (in years)","Unemployment \n (in %)", "region")
ss <- subset %>% select(-region)
c <- cor(ss)
smp_ss <- subset %>% filter(region == "Sub-Saharan Africa") %>% select(-region)
smp_eu <- subset %>% filter(region == "Europe & Central Asia") %>% select(-region)
smp_us <- subset %>% filter(region == "North America") %>% select(-region)

cc <- cor(smp_ss)
ceu <- cor(smp_eu)
cus <- cor(smp_us)

par(mfrow=c(2,2)) 
corrplot.mixed(c, tl.cex=0.7, tl.col="black")
mtext("Pearson Correlations for all Regions", at=2, line=1.5, cex=.9)
corrplot.mixed(cc, tl.cex=0.7, tl.col="black")
mtext("Pearson Correlations for Sub-Saharan Africa", at=2, line=1.5, cex=.9)
corrplot.mixed(ceu, tl.cex=0.7, tl.col="black")
mtext("Pearson Correlations for Europe & Central Asia", at=2, line=1.5, cex=.9)
corrplot.mixed(cus, tl.cex=0.7, tl.col="black")
mtext("Pearson Correlations for North America", at=2, line=1.5, cex=.9)
```

### ... and regression table...

``` {r RegTable1, results="asis",  size = "small"}
tab <- tab_regression_region$table
tab[12] <- paste(
  "\\\\[-1.8ex] & \\multicolumn{4}{c}{", smp_def$label[9], "} \\\\"
)
esc_var_names <- str_replace_all(smp_def$var_name, fixed ("_"), "\\_")
reg_names <- tibble(
  var_name = c(esc_var_names[c(10, 8)], "regionEurope  Central Asia", "regionLatin America  Caribbean" , "regionMiddle East  North Africa", "regionSouth Asia", "regionNorth America",
               "regionSub-Saharan Africa", "ln\\_gdp\\_capita:regionEurope  Central Asia", "ln\\_gdp\\_capita:regionLatin America  Caribbean", "ln\\_gdp\\_capita:regionMiddle East  North Africa",
               "ln\\_gdp\\_capita:regionSouth Asia", "ln\\_gdp\\_capita:regionNorth America", "ln\\_gdp\\_capita:regionSub-Saharan Africa"),
  label = c(smp_def$label[c(10, 8)], "Europe and Central Asia", "Latin America and Caribbean" , "Middle East and North Africa", "South Asia", "North America", "Sub-Saharan Africa", 
            "ln(GDP)*Europe and Central Asia", "ln(GDP)*Latin America and Caribbean", "ln(GDP)*Middle East and North Africa",
            "ln(GDP)*South Asia", "ln(GDP)*North America", "ln(GDP)*Sub-Saharan Africa")
)
for (i in 1:55) {
  for(j in 1:length(reg_names$label))
  tab[i] <- str_replace(
    tab[i], fixed(reg_names$var_name[j]), reg_names$label[j]
  )
}

latex_tab <- c(
  "\\begin{table}[!htbp]",
  "\\tiny",
  "\\setlength\\tabcolsep{1.5pt}",
  "\\caption{\\label{tab:reg}Regressions Part 1}",
  "\\centering",
  "\\begin{tabular}{lcccc}",
  tab[8:41],
  "\\end{tabular}",
  "\\end{table}"
)
# latex_tab[34] <- " \\bfseries \\color{red} Sub-Saharan Africa & \\bfseries \\color{red} $-$10.811$^{***}$ & \\bfseries \\color{red} $-$10.849$^{***}$ & \\bfseries \\color{red} $-$11.226$^{***}$ &  \\\\ "
latex_tab[35] <- " \\bfseries Sub-Saharan Africa & \\bfseries $-$10.811$^{***}$ & \\bfseries $-$10.849$^{***}$ & \\bfseries $-$11.226$^{***}$ &  \\\\ "
cat(paste(latex_tab, collapse = "\n"))
```

### ... continued

``` {r RegTable2, results="asis",  size = "small"}
tab <- tab_regression_region$table
tab[12] <- paste(
  "\\\\[-1.8ex] & \\multicolumn{4}{c}{", smp_def$label[9], "} \\\\"
)
esc_var_names <- str_replace_all(smp_def$var_name, fixed ("_"), "\\_")
reg_names <- tibble(
  var_name = c(esc_var_names[c(10, 8)], "regionEurope  Central Asia", "regionLatin America  Caribbean" , "regionMiddle East  North Africa", "regionSouth Asia", "regionNorth America",
               "regionSub-Saharan Africa", "ln\\_gdp\\_capita:regionEurope  Central Asia", "ln\\_gdp\\_capita:regionLatin America  Caribbean", "ln\\_gdp\\_capita:regionMiddle East  North Africa",
               "ln\\_gdp\\_capita:regionSouth Asia", "ln\\_gdp\\_capita:regionNorth America", "ln\\_gdp\\_capita:regionSub-Saharan Africa"),
  label = c(smp_def$label[c(10, 8)], "Europe and Central Asia", "Latin America and Caribbean" , "Middle East and North Africa", "South Asia", "North America", "Sub-Saharan Africa", 
            "ln(GDP)*Europe and Central Asia", "ln(GDP)*Latin America and Caribbean", "ln(GDP)*Middle East and North Africa",
            "ln(GDP)*South Asia", "ln(GDP)*North America", "ln(GDP)*Sub-Saharan Africa")
)
for (i in 1:55) {
  for(j in 1:length(reg_names$label))
  tab[i] <- str_replace(
    tab[i], fixed(reg_names$var_name[j]), reg_names$label[j]
  )
}

latex_tab <- c(
  "\\begin{table}[!htbp]",
  "\\tiny",
  "\\setlength\\tabcolsep{1.5pt}",
  "\\caption{\\label{tab:reg}Regressions Part 2}",
  "\\centering",
  "\\begin{tabular}{lcccc}",
  tab[42:69],
  "\\end{tabular}",
  "\\begin{tablenotes}[flushleft]",
  "\\itshape",
  "\\footnotetext",
  "\\item [Note: The reference region for the regional results of the regression output is: East Asia and Pacific. 
  \\newline Accordingly, the interaction-effects have to be interpreted relatively to the reference category.]",
  "\\end{tablenotes}",
  "\\end{table}"
)
# latex_tab[34] <- " \\bfseries \\color{red} Sub-Saharan Africa & \\bfseries \\color{red} $-$10.811$^{***}$ & \\bfseries \\color{red} $-$10.849$^{***}$ & \\bfseries \\color{red} $-$11.226$^{***}$ &  \\\\ "
# latex_tab[5] <- " \\bfseries Sub-Saharan Africa & \\bfseries $-$10.811$^{***}$ & \\bfseries $-$10.849$^{***}$ & \\bfseries $-$11.226$^{***}$ &  \\\\ "
cat(paste(latex_tab, collapse = "\n"))
```


### Conclusion

- The Preston Curve is alive and well! 
- Have fun adjusting this template to fit your research needs
- Consider contributing to its repository and promote Open Science!







